---
title: "Getting fitter"
author: "Luc Bams"
data: "May 9, 2019"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
    number_sections: true
---
```{r include=FALSE}
# Make sure the most recent data is loaded
source('~/R/Projects/fitbit/scripts/data_collection.R', encoding = 'UTF-8')
```

\*insert introductory section here\*  
\*make sure to come up with better titles for everything\*  

# Data
\*describe how the data were collected, stored, and what they look like\*

## Activity
\*describe activity data here\*

### Steps
```{r}
head(act_data$steps)
```

## Heart-Rate
\*describe heart-rate data here\*
```{r}
head(hr_intraday)
```

```{r}
hr_intraday %>%  
  summarise(
    total = sum(hr),
    minimum = min(hr),
    mean = mean(hr),
    median = median(hr),
    maximum = max(hr)
  ) %>% 
  knitr::kable(digits = 2, caption = "Summary of heart-rate data")
```

## Weight
\*describe weight data here\*

# How does getting fitter reflect in the data?  
\*obviously do a solid attempt to answer this question here\*  

## Weekly heart-rate distribution  
\*show ggridges weekly heart rate distributions here\*  

## Link steps/minute to bpm  
\*idea here is that when fitter heart rate is lower for a fixed amount of steps/min than when less fit\*  