---
title: "Taking Charge with Fitbit"
author: 
  - name: "Luc Bams"
    email: "luc-bams@hotmail.com"
    url: "https://github.com/Luc-Bams"
date: "Last Updated: `r format(Sys.time(), '%Y-%m-%d at %H:%M')`"
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: espresso
    number_sections: true
    code_folding: "hide"
---

```{r, data_collection_and_preparation, include = FALSE}
library(data.table)
library(ggridges)
library(scales)

# Make sure the most recent data is loaded
source('~/R/Projects/fitbit/scripts/data_collection.R', encoding = 'UTF-8')
```

\*insert introductory section here --> What is this? What motivated me?\*  
This document is the product of a combination of events in my life and some newly-fueled interest that I've had for a longer time.

  * Something about my interest in data:  
  
    + Gathering data on sports, physical state, games  
    + Bachelor Thesis about quantified self  
    
I'm fascinated by the all the cool things you can do with data, but I never really took the time to do something like it myself.   
\*make sure to come up with better titles for everything\*  

# Fitbit Time-Series Data
\*describe how the data were collected, stored, and what they look like\*  
All the data that I'm using can be retrieved by calling the fitbit API through the [getActivitiesResourceByDatePeriod](https://dev.fitbit.com/build/reference/web-api/explore/#/Time_Series/getActivitiesResourceByDatePeriod), [getHeartByDateIntraday](https://dev.fitbit.com/build/reference/web-api/explore/#/Heart_Rate_Intraday_Time_Series/getHeartByDateIntraday), and [getWeightByDate](https://dev.fitbit.com/build/reference/web-api/explore/#/Body_and_Weight/getWeightByDate) methods. The functions I wrote to perform these GET requests can be found in [this script](https://github.com/Luc-Bams/fitbit/blob/master/scripts/data_collection_functions.R). \*How I structured the data and why. --> leads to a reference to the script I used.\* The code to collect new data from the fitbit servers and tidy it can be found in [this script](https://github.com/Luc-Bams/fitbit/blob/master/scripts/data_collection.R).

* Minute-level
    + Continuous
        - calories
        - distance
        - elevation
    + Integer
        - floors
        - heartrate
        - steps
    + Categorical
        - activity intensity
* Daily-level
    + Continuous
        - weight
        - BMI
    + Integer
        - activity calories
        - base metabolic heartrate calories
        - rest heartrate
        - minutes (per heartrate zone --> separate table) [backlog]
        - calories out (per heartrate zone --> separate table) [backlog]

## Minute-Level Data

### Descriptive Statistics
```{r, ds_histogram_heartrate, fig.cap = "ds_histogram_heartrate"}
min_hr <- fitbit_data$ml_int %>%
  filter(type == "heartrate") %>%
  select(value) %>%
  min()

max_hr <- fitbit_data$ml_int %>%
  filter(type == "heartrate") %>%
  select(value) %>%
  max()

fitbit_data$ml_int %>%
  filter(type == "heartrate") %>%
  ggplot(aes(x = value, fill = value, stat(density))) +
  geom_histogram(binwidth = 1, color = "black", fill = "#2A211C") +
  geom_density(alpha = 0.8, color = "black", fill = "#9F2042") +
  theme_classic() +
  labs(x = "Heart rate", y = "Density") +
  scale_x_continuous(breaks = c(min_hr, 50, 75, 100, 125, 150, 175, max_hr))
```

```{r, ds_boxplot_heartrate, fig.cap = "ds_boxplot_heartrate"}
freq_total <- fitbit_data$ml_int %>%
  filter(type == "heartrate") %>%
  nrow()

fitbit_data$ml_int %>%
  filter(type == "heartrate") %>% 
  group_by(value) %>% 
  mutate(density = n()) %>%
  ungroup() %>%
  mutate(density = density/freq_total) %>%
  ggplot(aes(x = type, y = value, color = density)) +
  theme_minimal() +
  scale_color_viridis_c(option = "D") +
  geom_jitter(alpha = 0.03, height = 0, width = 0.6) +
  geom_violin(alpha = 0, position = position_dodge(width = .75), size = 1, color = "black") +
  geom_boxplot(notch = FALSE,  outlier.size = -1, color = "black", lwd = 1, alpha = 0, width = 0.04)

remove(freq_total)
```

```{r}
fitbit_data$ml_cont %>%
  group_by(type) %>%
  summarise(
  total = sum(value),
  minimum = min(value),
  mean = mean(value),
  median = median(value),
  maximum = max(value)
  ) %>%
  knitr::kable(digits = 2, caption = "Summary of activity data")
```

## Daily-Level Data
```{r}
# act_tidy %>% 
#   filter(!is.na(act_tidy$datetime)) %>%
#   group_by(as.Date(datetime, tz = "Europe/Amsterdam"), type) %>%
#   summarise(
#     daily_total = sum(value)
#   ) %>%
#   group_by(type) %>%
#   summarise(
#     total = sum(daily_total),
#     minimum = min(daily_total),
#     mean = mean(daily_total),
#     median = median(daily_total),
#     maximum = max(daily_total)
#   ) %>% 
#   knitr::kable(digits = 2, caption = "Summary of activity data")
```

```{r}
# act_tidy %>% 
#   filter(!is.na(act_tidy$datetime)) %>%
#   group_by(week, type) %>%
#   summarise(
#     average_day = sum(value)/uniqueN(day)
#   ) %>%
#   ggplot(aes(x = week, y = average_day)) +
#   geom_point() +
#   geom_line() +
#   facet_wrap(~ type, nrow = 2, ncol = 3, scales = "free_y")
```

# How does getting fitter reflect in the data?  
\*obviously do a solid attempt to answer this question here\*  

## Rest Heart Rate Over Time
```{r message=FALSE}
# hr_summary %>%
#   ggplot(aes(x = date, y = rest_hr)) +
#   geom_point() +
#   geom_smooth(se = FALSE, span = 0.275) +
#   labs(title = "Rest Heart Rate over Time",
#        subtitle = "*this is how you make a subtitle, Luc*")
```

```{r, weekly_heart_rate_summary_stats_plot, fig.cap = "weekly_heart_rate_summary_stats_plot"}
# hr_intraday %>%  
#   group_by(week) %>%
#   summarise(
#     minimum = min(hr),
#     mean = mean(hr),
#     median = median(hr),
#     maximum = max(hr)
#   ) %>% 
#   gather(key = "type", value = "value", -week) %>%
#   ggplot(aes(x = week, y = value)) +
#   geom_point() +
#   geom_line() +
#   facet_wrap(~ type, nrow = 2, ncol = 2, scales = "free_y")
```

## Weekly heart-rate distribution  
\*show ggridges weekly heart rate distributions here\* 
```{r, weekly_heart_rate_distribution_density_plot, fig.cap = "weekly_heart_rate_distribution_density_plot"}
# hr_intraday %>%
#   mutate(week = fct_rev(as.factor(week))) %>%
#   ggplot(aes(x= hr, y = week)) +
#   geom_density_ridges(scale = 2.5, alpha = 0.7) +
#   xlim(30, 194) +
#   theme_ridges()
```

## Link steps/minute to bpm  
\*idea here is that when fitter heart rate is lower for a fixed amount of steps/min than when less fit\*  

# Useful Resources

* [Fitbit Web API Basics](https://dev.fitbit.com/build/reference/web-api/basics/)
* [Fitbit Web API Swagger UI](https://dev.fitbit.com/build/reference/web-api/explore/#/)
* [Strava Web API Swagger UI](https://developers.strava.com/playground/#/)